---
title: "BIOCARD Data Analysis - CHAMPS"
author: "Yuhan Xiao, Christine Hou"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = NA,
                      fig.align = "center", fig.height = 6, fig.width = 12)
library(readxl)
library(tidyverse)
library(mice)
library(knitr)
library(here)
library(kableExtra)
library(gridExtra)
library(ggplot2)
library(cowplot)
library(lmerTest) # mixed linear model
```

# Descriptive Statistics

```{r}
data_champs_clean <- read.csv(here("processed-data/CHAMPS_clean_data.csv"), 
                              row.names = 1, check.names = FALSE)

# only contain baseline info for each patient
# calculate follow up time 
data_champs_clean_bl <- data_champs_clean %>%
  group_by(SUBJECT_ID) %>%
  mutate(follow_up_time = max(AgeAtVisit) - Age_CHAMPS) %>%
  fill(follow_up_time, .direction = "downup") %>%
  mutate(n_ntk = n()) %>%
  slice_head(n = 1) %>%
  ungroup()

summary_stats <- data.frame(
  # Total number of participants
  N = nrow(data_champs_clean_bl),

  # Number with a diagnosis of MCI or dementia at last visit
  N_with_MCI_dementia = sum(data_champs_clean$fup_Dx == 1, na.rm = TRUE),
  
  # Number of participants with 'Subject_ID' ≥ 400
  N_participants_400 = sum(data_champs_clean_bl$SUBJECT_ID >= 400, na.rm = TRUE),
  
  # Mean (SD) Age_CHAMPS
  Mean_Age_CHAMPS = sprintf("%.2f (%.2f)", mean(data_champs_clean_bl$Age_CHAMPS, na.rm = TRUE), sd(data_champs_clean_bl$Age_CHAMPS, na.rm = TRUE)),
  
  # Range of Age_CHAMPS
  Range_Age_CHAMPS = sprintf("(%.2f, %.2f)", min(data_champs_clean_bl$Age_CHAMPS, na.rm = TRUE), max(data_champs_clean_bl$Age_CHAMPS, na.rm = TRUE)),
  
  # N (%) female sex
  N_female = sum(data_champs_clean_bl$Sex_F == 1, na.rm = TRUE),
  Percent_female = sprintf("%.2f%%", (sum(data_champs_clean_bl$Sex_F == 1, na.rm = TRUE) / nrow(data_champs_clean_bl)) * 100),
  
  # Mean (SD) years of education
  Mean_EDUC = sprintf("%.2f (%.2f)", mean(data_champs_clean_bl$EDUC, na.rm = TRUE), sd(data_champs_clean_bl$EDUC, na.rm = TRUE)),
  
  # N (%) White race
  N_white = sum(data_champs_clean_bl$Race_White == 1, na.rm = TRUE),
  Percent_white = sprintf("%.2f%%", (sum(data_champs_clean_bl$Race_White == 1, na.rm = TRUE) / nrow(data_champs_clean_bl)) * 100),
  
  # N (%) APOE4 carriers
  N_APOE4 = sum(data_champs_clean_bl$APOE4 == 1, na.rm = TRUE),
  Percent_APOE4 = sprintf("%.2f%%", (sum(data_champs_clean_bl$APOE4 == 1, na.rm = TRUE) / nrow(data_champs_clean_bl)) * 100),
  
  # Mean (SD) years of follow-up
  Mean_Followup = sprintf("%.2f (%.2f)", mean(data_champs_clean_bl$follow_up_time, na.rm = TRUE), sd(data_champs_clean_bl$follow_up_time, na.rm = TRUE)),
  
  # Range of years of follow-up
  Range_Followup = sprintf("(%.2f, %.2f)", min(data_champs_clean_bl$follow_up_time, na.rm = TRUE), max(data_champs_clean_bl$follow_up_time, na.rm = TRUE)),
  
  # Mean (SD) number of blood biomarker measures
  Mean_Biomarkers = sprintf("%.2f (%.2f)", mean(data_champs_clean_bl$n_ntk, na.rm = TRUE), sd(data_champs_clean_bl$n_ntk, na.rm = TRUE)),
  
  # Range of number of blood biomarker measures
  Range_Biomarkers = sprintf("(%.2f, %.2f)", min(data_champs_clean_bl$n_ntk, na.rm = TRUE), max(data_champs_clean_bl$n_ntk, na.rm = TRUE)),
  
  # Mean (SD) time (in years) 
  Mean_Time = sprintf("%.2f (%.2f)", mean(data_champs_clean_bl$Age_NTK_bl - data_champs_clean_bl$Age_CHAMPS, na.rm = TRUE), sd(data_champs_clean_bl$Age_NTK_bl - data_champs_clean_bl$Age_CHAMPS, na.rm = TRUE)),
  
  # Mean Range (min, max) time (in years) 
  Range_Time = sprintf("(%.2f, %.2f)", min(data_champs_clean_bl$Age_NTK_bl - data_champs_clean_bl$Age_CHAMPS, na.rm = TRUE), max(data_champs_clean_bl$Age_NTK_bl - data_champs_clean_bl$Age_CHAMPS, na.rm = TRUE)),
  
  # Mean (SD) LOW_INT_FREQ
  Mean_LOW_INT = sprintf("%.2f (%.2f)", mean(data_champs_clean_bl$LOW_INT_FREQ, na.rm = TRUE), sd(data_champs_clean_bl$LOW_INT_FREQ, na.rm = TRUE)),
  
  # Mean (SD) HI_INT_FREQ
  Mean_HI_INT = sprintf("%.2f (%.2f)", mean(data_champs_clean_bl$HIGH_INT_FREQ, na.rm = TRUE), sd(data_champs_clean_bl$HIGH_INT_FREQ, na.rm = TRUE)),
  
  # Mean (SD) ALL_INT_FREQ
  Mean_ALL_INT = sprintf("%.2f (%.2f)", mean(data_champs_clean_bl$ALL_INT_FREQ, na.rm = TRUE), sd(data_champs_clean_bl$ALL_INT_FREQ, na.rm = TRUE)),
  
  # Mean (SD) BMI
  Mean_BMI = sprintf("%.2f (%.2f)", mean(data_champs_clean_bl$BMI, na.rm = TRUE), sd(data_champs_clean_bl$BMI, na.rm = TRUE)),
  
  # Mean (SD) GFAP at baseline CHAMPS
  Mean_GFAP = sprintf("%.2f (%.2f)", mean(data_champs_clean_bl[data_champs_clean_bl$GFAP_outlier==0,]$GFAP, na.rm = TRUE), sd(data_champs_clean_bl[data_champs_clean_bl$GFAP_outlier==0,]$GFAP, na.rm = TRUE)),
  
  # Mean (SD) NFL at baseline CHAMPS
  Mean_NFL = sprintf("%.2f (%.2f)", mean(data_champs_clean_bl[data_champs_clean_bl$NFL_outlier==0,]$NFL, na.rm = TRUE), sd(data_champs_clean_bl[data_champs_clean_bl$NFL_outlier==0,]$NFL, na.rm = TRUE)),
  
  # Mean (SD) PTAU181 at baseline CHAMPS
  Mean_PTAU181 = sprintf("%.2f (%.2f)", mean(data_champs_clean_bl[data_champs_clean_bl$PTAU181_outlier==0,]$PTAU181, na.rm = TRUE), sd(data_champs_clean_bl[data_champs_clean_bl$PTAU181_outlier==0,]$PTAU181, na.rm = TRUE)),
  
  # Mean (SD) AB42AB40 at baseline CHAMPS
  Mean_AB42AB40 = sprintf("%.2f (%.2f)", mean(data_champs_clean_bl[data_champs_clean_bl$AB42AB40_outlier==0,]$AB42AB40, na.rm = TRUE), sd(data_champs_clean_bl[data_champs_clean_bl$AB42AB40_outlier==0,]$AB42AB40, na.rm = TRUE)),
  
  # Mean (SD) PTAU181_AB42AB40 at baseline CHAMPS
  Mean_PTAU181_AB42AB40 = sprintf("%.2f (%.2f)", mean(data_champs_clean_bl[data_champs_clean_bl$AB42AB40_outlier==0 & data_champs_clean_bl$PTAU181_outlier==0,]$PTAU181_AB42AB40, na.rm = TRUE), sd(data_champs_clean_bl[data_champs_clean_bl$AB42AB40_outlier==0 & data_champs_clean_bl$PTAU181_outlier==0,]$PTAU181_AB42AB40, na.rm = TRUE)),
  
  # Mean (SD) sTREM2 at baseline CHAMPS
  Mean_sTREM2 = sprintf("%.2f (%.2f)", mean(data_champs_clean_bl[data_champs_clean_bl$sTREM2_outlier==0,]$sTREM2, na.rm = TRUE), sd(data_champs_clean_bl[data_champs_clean_bl$sTREM2_outlier==0,]$sTREM2, na.rm = TRUE)),
  
  # Mean (SD) YKL40 at baseline CHAMPS
  Mean_YKL40 = sprintf("%.2f (%.2f)", mean(data_champs_clean_bl[data_champs_clean_bl$YKL40_outlier==0,]$YKL40, na.rm = TRUE), sd(data_champs_clean_bl[data_champs_clean_bl$YKL40_outlier==0,]$YKL40, na.rm = TRUE))
)
```

```{r, echo=FALSE}
summary_stats <- as.data.frame(t(summary_stats))
summary_stats$Variable = c("N", 
                           "N with a diagnosis of MCI or dementia at last visit (‘fup_Dx’ = 1)",
                           "N participants with ‘Subject_ID’ ≥ 400 *",
                           "Mean (SD) Age_CHAMPS", 
                           "Range (min, max) Age_CHAMPS", 
                           "N female sex (‘Sex_F’ = 1)", 
                           "% female sex (‘Sex_F’ = 1)", 
                           "Mean (SD) years of education (‘EDUC’) ",
                           "N White race (‘Race_White’ = 1)", 
                           "% White race (‘Race_White’ = 1)", 
                           "N APOE4 carriers (‘APOE4’ = 1)",
                           "% APOE4 carriers (‘APOE4’ = 1)",
                           "Mean (SD) years of follow-up (baseline CHAMPS to last blood biomarker measure)",
                           "Range (min, max) years of follow-up (baseline CHAMPS to last blood biomarker measure)", 
                           "Mean (SD) number of blood biomarker measures over time", 
                           "Range (min, max) number of blood biomarker measures over time",
                           "Mean (SD) time (in years) between baseline CHAMPS and associated baseline blood biomarker measure",
                           "Range (min, max) time (in years) between baseline CHAMPS and associated baseline blood biomarker measure", 
                           "Mean (SD) LOW_INT_FREQ", 
                           "Mean (SD) HI_INT_FREQ",
                           "Mean (SD) ALL_INT_FREQ",
                           "Mean (SD) BMI", 
                           "Mean (SD) GFAP at baseline actigraphy", 
                           "Mean (SD) NFL at baseline actigraphy", 
                           "Mean (SD) PTAU181 at baseline actigraphy", 
                           "Mean (SD) AB42AB40 at baseline actigraphy", 
                           "Mean (SD) PTAU181_AB42AB40 at baseline actigraphy", 
                           "Mean (SD) sTREM2 at baseline actigraphy", 
                           "Mean (SD) YKL40 at baseline actigraphy"
)
colnames(summary_stats) = c("Participants in analyses", "")

summary_stats %>%
  kable(caption = "Descriptive Statistics for Participants Included in the CHAMPS Analyses")
```


# Model Analysis

### Model Specification

The linear mixed-effects model used in this analysis is specified as follows:

$$
Y_{ij} = \beta_0 + \beta_1 \text{time}_{ij} + \beta_2 \text{Age_CHAMPS}_i + \beta_3 \text{Sex_F}_i + \beta_4 \text{APOE4}_i + \beta_5 \text{BMI}_i + \beta_6 \text{CHAMPS}_i
+ \\ \beta_7 (\text{time}_{ij} \times \text{Age_CHAMPS}_i) + \beta_8 (\text{time}_{ij} \times \text{Sex_F}_i) + \beta_9 (\text{time}_{ij} \times \text{APOE4}_i) + \beta_{10} (\text{time}_{ij} \times \text{BMI}_i) + \beta_{11} (\text{time}_{ij} \times \text{CHAMPS}_i) \\
+ u_{0i} + u_{1i} \text{time}_{ij} + \epsilon_{ij}
$$

Where:

- $Y_{ij}$ is the biomarker outcome for participant $i$ at time $j$. 
- $\beta_0$ is the fixed intercept. 
- $\beta_1, \dots, \beta_{11}$ are the fixed effect coefficients. 
- $u_{0i} \sim N(0, \sigma^2_u)$ is the random intercept for subject $i$. 
- $u_{1i} \sim N(0, \sigma^2_v)$ is the random slope for time for subject $i$. 
- $\epsilon_{ij} \sim N(0, \sigma^2)$ is the residual error.

### Data Preparation

```{r}
model_vars = c("Age_CHAMPS_z", "bl_sex", "bl_APOE4", "bl_BMI_z", "bl_low_z", 
               "bl_high_z","bl_all_z","time")

# mutate baseline characteristics
data_champs_model <- data_champs_clean %>%
  mutate(time = AgeAtVisit - Age_CHAMPS) %>%
  arrange(SUBJECT_ID, VISITNO) %>%
  group_by(SUBJECT_ID) %>%
  mutate(bl_sex = ifelse(first_champs_visit == 1, Sex_F, NA),
         bl_APOE4 = ifelse(first_champs_visit == 1, APOE4, NA),
         bl_BMI = ifelse(first_champs_visit == 1, BMI, NA),
         bl_low = ifelse(first_champs_visit == 1, LOW_INT_FREQ, NA),
         bl_high = ifelse(first_champs_visit == 1, HIGH_INT_FREQ, NA),
         bl_all = ifelse(first_champs_visit == 1, ALL_INT_FREQ, NA)
         ) %>%
  # standardize continuous predictors
  mutate(bl_sex = as.factor(bl_sex),
         bl_APOE4 = as.factor(bl_APOE4),
         Age_CHAMPS_z = (Age_CHAMPS - mean(data_champs_clean_bl$Age_CHAMPS, na.rm = TRUE)) / sd(data_champs_clean_bl$Age_CHAMPS, na.rm = TRUE),
         bl_BMI_z = (bl_BMI - mean(data_champs_clean_bl$BMI, na.rm = TRUE)) / sd(data_champs_clean_bl$BMI, na.rm = TRUE),
         bl_low_z = (bl_low - mean(data_champs_clean_bl$LOW_INT_FREQ, na.rm = TRUE)) / sd(data_champs_clean_bl$LOW_INT_FREQ, na.rm = TRUE),
         bl_high_z = (bl_high - mean(data_champs_clean_bl$HIGH_INT_FREQ, na.rm = TRUE)) / sd(data_champs_clean_bl$HIGH_INT_FREQ, na.rm = TRUE),
         bl_all_z = (bl_all - mean(data_champs_clean_bl$ALL_INT_FREQ, na.rm = TRUE)) / sd(data_champs_clean_bl$ALL_INT_FREQ, na.rm = TRUE)
         ) %>%
  fill(bl_sex, .direction = "down") %>%
  fill(bl_APOE4, .direction = "down") %>%
  fill(Age_CHAMPS_z, .direction = "down") %>%
  fill(bl_BMI_z, .direction = "down") %>%
  fill(bl_low_z, .direction = "down") %>%
  fill(bl_high_z, .direction = "down") %>%
  fill(bl_all_z, .direction = "down") 
# 9 rows missing BMI

data_champs_GFAP <- data_champs_model %>%
  select(c(SUBJECT_ID, model_vars, 
           GFAP_zscore, GFAP_outlier)) %>%
  filter(GFAP_outlier == 0) %>%
  select(-c(GFAP_outlier))

data_champs_NFL <- data_champs_model %>%
  select(c(SUBJECT_ID, model_vars, 
           NFL_zscore, NFL_outlier)) %>%
  filter(NFL_outlier == 0) %>%
  select(-c(NFL_outlier))

data_champs_PTAU181 <- data_champs_model %>%
  select(c(SUBJECT_ID, model_vars, 
           PTAU181_zscore, PTAU181_outlier)) %>%
  filter(PTAU181_outlier == 0) %>%
  select(-c(PTAU181_outlier))

data_champs_AB42AB40 <- data_champs_model %>%
  select(c(SUBJECT_ID, model_vars, 
           AB42AB40_zscore, AB42AB40_outlier)) %>%
  filter(AB42AB40_outlier == 0) %>%
  select(-c(AB42AB40_outlier))

data_champs_sTREM2 <- data_champs_model %>%
  select(c(SUBJECT_ID, model_vars, 
           sTREM2_zscore, sTREM2_outlier)) %>%
  filter(sTREM2_outlier == 0) %>%
  select(-c(sTREM2_outlier))

data_champs_YKL40 <- data_champs_model %>%
  select(c(SUBJECT_ID, model_vars, 
           YKL40_zscore, YKL40_outlier)) %>%
  filter(YKL40_outlier == 0) %>%
  select(-c(YKL40_outlier))
```

### Low Intensity

#### GFAP

```{r GFAP low}
model_GFAP_low <- lmerTest::lmer(GFAP_zscore ~ time * (Age_CHAMPS_z + bl_sex + bl_APOE4 + bl_BMI_z + bl_low_z) + (1 + time | SUBJECT_ID), data = data_champs_GFAP)
summary(model_GFAP_low)
```

#### NFL

```{r NFL low}
model_NFL_low <- lmerTest::lmer(NFL_zscore ~ time * (Age_CHAMPS_z + bl_sex + bl_APOE4 + bl_BMI_z + bl_low_z) + (1 + time | SUBJECT_ID), data = data_champs_NFL)
summary(model_NFL_low)
```

#### PTAU181

```{r PTAU181 low}
model_PTAU181_low <- lmerTest::lmer(PTAU181_zscore ~ time * (Age_CHAMPS_z + bl_sex + bl_APOE4 + bl_BMI_z + bl_low_z) + (1 + time | SUBJECT_ID), data = data_champs_PTAU181)
summary(model_PTAU181_low)
```

In PTAU181 analysis, low intensity CHAMPS value is a predictor with a significance of 0.05.

```{r PTAU181 low visual}
median_int <- median(data_champs_PTAU181$bl_low_z, na.rm = TRUE)

data_champs_PTAU181$int_group <- ifelse(data_champs_PTAU181$bl_low_z > median_int, "Above Median", "Below Median")

plot_below <- ggplot(data_champs_PTAU181[data_champs_PTAU181$int_group == "Below Median",], aes(x = time, y = PTAU181_zscore)) +
  geom_point(color = "blue") +
  geom_smooth(color = "blue") +
  labs(title = "Standardized Biomarker PTAU181 Values over Time \nSince Baseline (Below Median Low Intensity Value)",
       x = "Time (years)",
       y = "Standardized PTAU181 Score") +
  theme_minimal()

plot_above <- ggplot(data_champs_PTAU181[data_champs_PTAU181$int_group == "Above Median",], aes(x = time, y = PTAU181_zscore)) +
  geom_point(color = "red") +
  geom_smooth(color = "red") +
  labs(title = "Standardized Biomarker PTAU181 Values over Time \nSince Baseline (Above Median Low Intensity Value)",
       x = "Time (years)",
       y = "Standardized PTAU181 Score") +
  theme_minimal()
grid.arrange(plot_below, plot_above, ncol = 2)
```

#### AB42AB40

```{r AB42AB40 low}
model_AB42AB40_low <- lmerTest::lmer(AB42AB40_zscore ~ time * (Age_CHAMPS_z + bl_sex + bl_APOE4 + bl_BMI_z + bl_low_z) + (1 + time | SUBJECT_ID), data = data_champs_AB42AB40)
summary(model_AB42AB40_low)
```

#### sTREM2

```{r sTREM2 low}
model_sTREM2_low <- lmerTest::lmer(sTREM2_zscore ~ time * (Age_CHAMPS_z + bl_sex + bl_APOE4 + bl_BMI_z + bl_low_z) + (1 + time | SUBJECT_ID), data = data_champs_sTREM2)
summary(model_sTREM2_low)
```

#### YKL40

```{r YKL40 low}
model_YKL40_low <- lmerTest::lmer(YKL40_zscore ~ time * (Age_CHAMPS_z + bl_sex + bl_APOE4 + bl_BMI_z + bl_low_z) + (1 + time | SUBJECT_ID), data = data_champs_YKL40)
summary(model_YKL40_low)
```

#### Summary

```{r low result extract}
extract_results <- function(model) {
  model_summary <- summary(model)
  
  estimate_main <- model_summary$coefficients[7, 1]
  std.error_main <- model_summary$coefficients[7, 2]
  t_value_main <- model_summary$coefficients[7, 4]
  p_value_main <- model_summary$coefficients[7, 5]
  
  estimate_interaction <- model_summary$coefficients[12, 1]
  std.error_interaction <- model_summary$coefficients[12, 2]
  t_value_interaction <- model_summary$coefficients[12, 4]
  p_value_interaction <- model_summary$coefficients[12, 5]
  
  result_main <- c(Estimate = estimate_main, Std.Error = std.error_main, T_value = t_value_main, P_value = p_value_main)
  result_interaction <- c(Estimate = estimate_interaction, Std.Error = std.error_interaction, T_value = t_value_interaction, P_value = p_value_interaction)
  
  return(rbind(result_main, result_interaction))
}

# Create a list of models
models <- list(
  GFAP = model_GFAP_low,
  NFL = model_NFL_low,
  PTAU181 = model_PTAU181_low,
  AB42AB40 = model_AB42AB40_low,
  sTREM2 = model_sTREM2_low,
  YKL40 = model_YKL40_low
)

results <- lapply(models, extract_results)
results_df <- do.call(rbind, results)
# Convert the row names (biomarker names) into a column
results_df <- data.frame(Biomarker = rep(names(models), each = 2), results_df, row.names = NULL)
results_df$Interaction <- rep(c("Low Intensity", "Low Intensity * Time"), length(models))
results_df <- as.data.frame(results_df[, c("Biomarker", "Interaction", "Estimate", "Std.Error", "T_value", "P_value")])

kable(results_df, col.names = c("Biomarker", "Variable", "Estimate", "Std.Error", "t value", "P-value"))  %>%
 kable_styling("striped", full_width = T) %>%
 column_spec(1, bold = T) %>%
 collapse_rows(columns = 1, valign = "middle") 
```

### High Intensity

#### GFAP

```{r GFAP high} 
model_GFAP_high <- lmerTest::lmer(GFAP_zscore ~ time * (Age_CHAMPS_z + bl_sex + bl_APOE4 + bl_BMI_z + bl_high_z) + (1 + time | SUBJECT_ID), data = data_champs_GFAP)
summary(model_GFAP_high)
```

#### NFL

```{r NFL high}
model_NFL_high <- lmerTest::lmer(NFL_zscore ~ time * (Age_CHAMPS_z + bl_sex + bl_APOE4 + bl_BMI_z + bl_high_z) + (1 + time | SUBJECT_ID), data = data_champs_NFL)
summary(model_NFL_high)
```

In NFL analysis, the interaction of high intensity CHAMPS value with time is a predictor with a significance of 0.01.

```{r NFl high visual}
median_int <- median(data_champs_NFL$bl_high_z, na.rm = TRUE)

data_champs_NFL$int_group <- ifelse(data_champs_NFL$bl_high_z > median_int, "Above Median", "Below Median")

plot_below <- ggplot(data_champs_NFL[data_champs_NFL$int_group == "Below Median",], aes(x = time, y = NFL_zscore)) +
  geom_point(color = "blue") +
  geom_smooth(color = "blue") +
  labs(title = "Standardized Biomarker NFL Values over Time \nSince Baseline (Below Median High Intensity Value)",
       x = "Time (years)",
       y = "Standardized NFL Score") +
  theme_minimal()

plot_above <- ggplot(data_champs_NFL[data_champs_NFL$int_group == "Above Median",], aes(x = time, y = NFL_zscore)) +
  geom_point(color = "red") +
  geom_smooth(color = "red") +
  labs(title = "Standardized Biomarker NFL Values over Time \nSince Baseline (Above Median High Intensity Value)",
       x = "Time (years)",
       y = "Standardized NFL Score") +
  theme_minimal()
grid.arrange(plot_below, plot_above, ncol = 2)
```

#### PTAU181

```{r PTAU181 high}
model_PTAU181_high <- lmerTest::lmer(PTAU181_zscore ~ time * (Age_CHAMPS_z + bl_sex + bl_APOE4 + bl_BMI_z + bl_high_z) + (1 + time | SUBJECT_ID), data = data_champs_PTAU181)
summary(model_PTAU181_high)
```

#### AB42AB40

```{r AB42AB40 high}
model_AB42AB40_high <- lmerTest::lmer(AB42AB40_zscore ~ time * (Age_CHAMPS_z + bl_sex + bl_APOE4 + bl_BMI_z + bl_high_z) + (1 + time | SUBJECT_ID), data = data_champs_AB42AB40)
summary(model_AB42AB40_high)
```

#### sTREM2

```{r sTREM2 high}
model_sTREM2_high <- lmerTest::lmer(sTREM2_zscore ~ time * (Age_CHAMPS_z + bl_sex + bl_APOE4 + bl_BMI_z + bl_high_z) + (1 + time | SUBJECT_ID), data = data_champs_sTREM2)
summary(model_sTREM2_high)
```

#### YKL40

```{r YKL40 high}
model_YKL40_high <- lmerTest::lmer(YKL40_zscore ~ time * (Age_CHAMPS_z + bl_sex + bl_APOE4 + bl_BMI_z+ bl_high_z) + (1 + time | SUBJECT_ID), data = data_champs_YKL40)
summary(model_YKL40_high)
```

#### Summary

```{r high result extract}
# Create a list of models
models <- list(
  GFAP = model_GFAP_high,
  NFL = model_NFL_high,
  PTAU181 = model_PTAU181_high,
  AB42AB40 = model_AB42AB40_high,
  sTREM2 = model_sTREM2_high,
  YKL40 = model_YKL40_high
)

results <- lapply(models, extract_results)
results_df <- do.call(rbind, results)
# Convert the row names (biomarker names) into a column
results_df <- data.frame(Biomarker = rep(names(models), each = 2), results_df, row.names = NULL)
results_df$Interaction <- rep(c("High Intensity", "High Intensity * Time"), length(models))
results_df <- as.data.frame(results_df[, c("Biomarker", "Interaction", "Estimate", "Std.Error", "T_value", "P_value")])

kable(results_df, col.names = c("Biomarker", "Variable", "Estimate", "Std.Error", "t value", "P-value"))  %>%
 kable_styling("striped", full_width = T) %>%
 column_spec(1, bold = T) %>%
 collapse_rows(columns = 1, valign = "middle") 
```

### All Intensity

#### GFAP

```{r GFAP all}
model_GFAP_all <- lmerTest::lmer(GFAP_zscore ~ time * (Age_CHAMPS_z + bl_sex + bl_APOE4 + bl_BMI_z + bl_all_z) + (1 + time | SUBJECT_ID), data = data_champs_GFAP)
summary(model_GFAP_all)
```

#### NFL

```{r NFL all}
model_NFL_all <- lmerTest::lmer(NFL_zscore ~ time * (Age_CHAMPS_z + bl_sex + bl_APOE4 + bl_BMI_z + bl_all_z) + (1 + time | SUBJECT_ID), data = data_champs_NFL)
summary(model_NFL_all)
```

In NFL analysis, the interaction of all intensity CHAMPS value with time is a predictor with a significance of 0.01.

```{r NFL all visual}
median_int <- median(data_champs_NFL$bl_all_z, na.rm = TRUE)

data_champs_NFL$int_group <- ifelse(data_champs_NFL$bl_all_z > median_int, "Above Median", "Below Median")

plot_below <- ggplot(data_champs_NFL[data_champs_NFL$int_group == "Below Median",], aes(x = time, y = NFL_zscore)) +
  geom_point(color = "blue") +
  geom_smooth(color = "blue") +
  labs(title = "Standardized Biomarker NFL Values over Time \nSince Baseline (Below Median All Intensity Value)",
       x = "Time (years)",
       y = "Standardized NFL Score") +
  theme_minimal()

plot_above <- ggplot(data_champs_NFL[data_champs_NFL$int_group == "Above Median",], aes(x = time, y = NFL_zscore)) +
  geom_point(color = "red") +
  geom_smooth(color = "red") +
  labs(title = "Standardized Biomarker NFL Values over Time \nSince Baseline (Above Median All Intensity Value)",
       x = "Time (years)",
       y = "Standardized NFL Score") +
  theme_minimal()
grid.arrange(plot_below, plot_above, ncol = 2)
```


#### PTAU181

```{r PTAU181 all}
model_PTAU181_all <- lmerTest::lmer(PTAU181_zscore ~ time * (Age_CHAMPS_z + bl_sex + bl_APOE4 + bl_BMI_z + bl_all_z) + (1 + time | SUBJECT_ID), data = data_champs_PTAU181)
summary(model_PTAU181_all)
```

#### AB42AB40

```{r AB42AB40 all}
model_AB42AB40_all <- lmerTest::lmer(AB42AB40_zscore ~ time * (Age_CHAMPS_z + bl_sex + bl_APOE4 + bl_BMI_z + bl_all_z) + (1 + time | SUBJECT_ID), data = data_champs_AB42AB40)
summary(model_AB42AB40_all)
```

#### sTREM2

```{r sTREM2 all}
model_sTREM2_all <- lmerTest::lmer(sTREM2_zscore ~ time * (Age_CHAMPS_z + bl_sex + bl_APOE4 + bl_BMI_z + bl_all_z) + (1 + time | SUBJECT_ID), data = data_champs_sTREM2)
summary(model_sTREM2_all)
```

#### YKL40

```{r YKL40 all}
model_YKL40_all <- lmerTest::lmer(YKL40_zscore ~ time * (Age_CHAMPS_z + bl_sex + bl_APOE4 + bl_BMI_z + bl_all_z) + (1 + time | SUBJECT_ID), data = data_champs_YKL40)
summary(model_YKL40_all)
```

#### Summary

```{r all result extract}
# Create a list of models
models <- list(
  GFAP = model_GFAP_all,
  NFL = model_NFL_all,
  PTAU181 = model_PTAU181_all,
  AB42AB40 = model_AB42AB40_all,
  sTREM2 = model_sTREM2_all,
  YKL40 = model_YKL40_all
)

results <- lapply(models, extract_results)
results_df <- do.call(rbind, results)
# Convert the row names (biomarker names) into a column
results_df <- data.frame(Biomarker = rep(names(models), each = 2), results_df, row.names = NULL)
results_df$Interaction <- rep(c("All Intensity", "All Intensity * Time"), length(models))
results_df <- as.data.frame(results_df[, c("Biomarker", "Interaction", "Estimate", "Std.Error", "T_value", "P_value")])

kable(results_df, col.names = c("Biomarker", "Variable", "Estimate", "Std.Error", "t value", "P-value"))  %>%
 kable_styling("striped", full_width = T) %>%
 column_spec(1, bold = T) %>%
 collapse_rows(columns = 1, valign = "middle") 
```
